---
title: "ROSMAP PiL DEG analysis"
output:
  html_notebook: default
  pdf_document: default
---

First we load in the PiL data


```{r}
foo <- data.table::fread('ROSMAP_PiL_dep_cog.csv',data.table=F)
foo <- dplyr::filter(foo,fu_year==0)
```

Next we load in the RNAseq design matrix

```{r}
synapser::synLogin()
dmobj <- synapser::synGet('syn8456640')
dm <- data.table::fread(dmobj$path,data.table=F)
```

Next we load in the covariate data

```{r}
covobj <- synapser::synGet('syn3191087')
cova <- data.table::fread(covobj$path,data.table=F)
cova <- cova[,-1]
colnames(cova)
cova <- dplyr::select(cova,projid,cogdx,braaksc,ceradsc,age_first_ad_dx,educ)
```

Next we load in the count data

```{r}
countobj <- synapser::synGet('syn8456638')
countMatrix <- data.table::fread(countobj$path,data.table=F)
rownames(countMatrix) <- countMatrix$ensembl_gene_id
countMatrix <- countMatrix[,-1]
countMatrix <- t(countMatrix)
countMatrix <- data.frame(countMatrix,stringsAsFactors = F)
countMatrix$sampleId <- rownames(countMatrix)
```

Now we join all of the relevant data frames together into a single analysis data frame

```{r}
key <- data.table::fread("key.csv",data.table=F)
foo <- dplyr::left_join(foo,key)
foo <- dplyr::filter(foo,!is.na(Sampleid))
foo <- dplyr::left_join(foo,dm,by=c('Sampleid'='SampleID'))
foo <- dplyr::left_join(foo,cova)
foo <- dplyr::left_join(foo,countMatrix,by=c('Sampleid'='sampleId'))
```

Now we remove individuals with dementia at baseline
```{r}
foo <- dplyr::filter(foo,dcfdx < 4)
```


Define the simple linear model function and get the gene names

```{r}
run_simple_model <- function(geneId,pheno,foo,pb,braak=FALSE){
  if(!braak){
    str <- paste0(pheno," ~ ",geneId," + dcfdx + Batch0 + Batch1 + Batch2 + Batch3 + Batch4 + Batch5 + Batch6 + Batch7 + msex0 + RINcontinuous + PCT_CODING_BASES + PCT_INTERGENIC_BASES + pmi + age_at_visit")
  }else{
    str <- paste0(pheno," ~ ",geneId," + dcfdx + Batch0 + Batch1 + Batch2 + Batch3 + Batch4 + Batch5 + Batch6 + Batch7 + msex0 + RINcontinuous + PCT_CODING_BASES + PCT_INTERGENIC_BASES + pmi + age_at_visit + braaksc")
  }
  str <- as.formula(str)
  pb$tick()
  #res<-lme4::lmer(str,foo)
  res <- lm(str,foo)
  return(summary(res)$coef[2,4])
}

gene_ids <- colnames(countMatrix)
gene_ids <- gene_ids[-which(gene_ids=='sampleId')]
```

Run model 1, educ

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'educ',foo,pb))
gap::qqunif(pil_lm_results,main='educ, Model 1')
```

Run model 2, PiL

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'educ',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='educ, Model 2')
```



Run model 1, PiL

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'purpose_total',foo,pb))
gap::qqunif(pil_lm_results,main='purpose_total, Model 1')
```

Run model 2, PiL

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'purpose_total',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='purpose_total, Model 2')
```

Run model 1, r_depres

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'r_depres',foo,pb))
gap::qqunif(pil_lm_results,main='r_depres, Model 1')
```

Run model 2, r_depres

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'r_depres',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='r_depres, Model 2')
```

Run model 1, cogn_ep

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_ep',foo,pb))
gap::qqunif(pil_lm_results,main='cogn_ep, Model 1')
```

Run model 2, cogn_ep

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_ep',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='cogn_ep, Model 2')
```

Run model 1, cogn_po

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_po',foo,pb))
gap::qqunif(pil_lm_results,main='cogn_po, Model 1')
```

Run model 2, cogn_po

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_po',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='cogn_po, Model 2')
```

Run model 1, cogn_ps

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_ps',foo,pb))
gap::qqunif(pil_lm_results,main='cogn_ps, Model 1')
```

Run model 2, cogn_ps

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_ps',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='cogn_ps, Model 2')
```

Run model 1, cogn_se

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_se',foo,pb))
gap::qqunif(pil_lm_results,main='cogn_se, Model 1')
```

Run model 2, cogn_se

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_se',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='cogn_se, Model 2')
```

Run model 1, cogn_wo

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_wo',foo,pb))
gap::qqunif(pil_lm_results,main='cogn_wo, Model 1')
```

Run model 2, cogn_wo

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_wo',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='cogn_wo, Model 2')
```

Run model 1, cogn_global

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_global',foo,pb))
gap::qqunif(pil_lm_results,main='cogn_global, Model 1')
```

Run model 2, cogn_global

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cogn_global',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='cogn_global, Model 2')
```

Run model 1, cesdsum

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cesdsum',foo,pb))
gap::qqunif(pil_lm_results,main='cesdsum, Model 1')
```

Run model 2, cesdsum

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'cesdsum',foo,pb,braak=TRUE))
gap::qqunif(pil_lm_results,main='cesdsum, Model 2')
```
Run model 1, braaksc

```{r}
pb <- progress::progress_bar$new(total=(length(gene_ids)))
system.time(pil_lm_results <- sapply(gene_ids,run_simple_model,'braaksc',foo,pb))
gap::qqunif(pil_lm_results,main='braaksc, Model 1')
```
